---
title: "Simulations"
date: "July 21, 2016"
---

```{r echo = FALSE, include = F, eval = T}
library(ggplot2)
library(dplyr)
```

## Motivation

* Probability theory is complimentary to statistical inference.
* A __statistic__ is often just a function of a random sample (i.e. sample mean, 95th percentile, sample proportion)
* Statistics are often used as __estimators__ of quantities of interest about the distribution, called __parameters__. Statistics are random variables (since they depend on the sample); parameters are fixed.
* In simple cases, we can study the __sampling distribution__ of the statistic analytically. But in more complicated cases, we turn to simulations.

---

## Simulations

* The main idea in a simulation study is to replace the mathematical expression for the distribution with a __sample__ from that distribution.


#### Example

* Suppose  

$$X_1, X_2, ..., X_n$$ 

are independent observations from the same distribution with expected value $\mu$ and standard deviation $\sigma$. 

  + We are interested in the distribution of $\bar{X}$. So we take many samples of size $n$, and study the behavior of the sample averages.
  
---  

## Simulations

* We use the sample to estimate features of the distribution, such as the behavior of various statistics under repeated sampling from the distribution. 
* The downside to using simulations is that while mathematical results are symbolic, in terms of arbitrary parameters and sample sizes, these values must be specified in a simulation.

---

## Steps in carrying out a simulation study

1. Specify what makes up an individual experiment: sample size, distributions, parameters, statistic of interest
2. Write an expression or function to carry out an individual experiment and return the statistic
3. Determine what inputs, if any, to vary (e.g. different sample sizes or parameters)
4. For each combination of inputs, repeat the experiment B times, providing B samples of the statistic.
5. For each combination of inputs, summarize the empirical distribution of the statistic of interest.

---

## Sampling

* `sample(x, size, replace = FALSE, prob = NULL)` takes a sample of a specified size from the elements of a vector `x` using either with or without replacement.
* Note: If your vector `x` is of length 1, then R will sample from `1:x`

```{r}
sample(10, size = 5)
```

---

## Examples 

* We can simulate rolling a die 5 times using:

```{r}
sample(1:6, size = 5, replace = TRUE)
```

* We can simulate tossing a biased coin 10 times using: 

```{r}
sample(c("H", "T"), size = 10, replace = TRUE, prob = c(7/10, 3/10))
```

---

## Replicate

* `replicate(n, expr)` allows us to repeat an experiment `n` times in order to obtain the empirical distribution of some statistic of interest

---

## Example

* Suppose we're interested in the distribution of the sum of 6 rolls of a fair die. 

```{r}
sumRolls <- function(x) {
  sample(1:6, size = x, replace = TRUE) %>% 
    sum()
}

set.seed(123456)
results <- replicate(10000, sumRolls(6))
head(results, 100)
```

---

* We know by the Central Limit Theorem that for large B, the distribution of the sum will be approximately normal. But we can also arrive at the same conclusion by making inferences from the empirical distribution.

```{r, fig.width = 8, fig.height = 4}
ggplot() +
  aes(x = results, y = ..density..) +
  geom_histogram(binwidth = 1) +
  geom_density() +
  scale_x_continuous(breaks = seq(6, 36, by = 6)) +
  labs(x = "Sum",
       y = "Density", 
       title = "Distribution of Sum of 6 Rolls of a Fair Die")
```

---

## Generating random samples from a distribution

* `runif(n, min = 0, max = 1)` samples from the uniform distribution on the interval [0,1].
* The chance the value drawn is:

  + between 0 and 1/3 is equal to 1/3
  + between 1/3 and 1/2 is equal to 1/6
  + between 9/10 and 1 is equal to 1/10
  
  
* `rnorm(n, mean = 0, sd = 1)` samples from the normal distribution with center = `mean` and spread = `sd`


* `rbinom(n, size, prob)` samples from the binomial distribution with number of trials = `size` and chance of success = `prob`

---

## Other Random Number Functions

Distribution     | Function
---------------- | -----------------------------------------------------
Geometric        | `rgeom(n, prob)`
Poisson          | `rpois(n, lambda)`
Exponential      | `rexp(n, rate = 1)`
Gamma            | `rgamma(n, shape, rate = 1, scale = 1/rate)`
t                | `rt(n, df)`
F                | `rf(n, df1, df2)`



* See `?Distributions` for others

---

## How does R generate random numbers?

* R uses a pseudo-random number generator:

  + It starts with a seed and an algorithm
  + The seed is plugged into the algorithm and a number is returned
  + That new number is then plugged into the algorithm, amd the next number is created
  
* The algorithms are created so that the numbers produced behave like random values

---

## Setting the seed

* One advantage to pseudo-random number generating is that you can reproduce simulation results by controlling the seed

```{r}
set.seed(123456)
rnorm(5)
```

```{r}
rnorm(5)
```

```{r}
set.seed(123456)
rnorm(5)
```



