---
title: Apply Functions
layout: default
output: html_document
---

```{r, echo = FALSE, message = FALSE}
library(pryr)
```

## Motivation

Imagine you've loaded a data file, like the one below, that uses $-99$ to represent missing values. You want to replace all the $-99$s with `NA`s.

```{r}
# Generate a sample dataset
set.seed(1014)
df <- data.frame(replicate(6, sample(c(1:10, -99), 6, rep = TRUE)))
names(df) <- letters[1:6]
df
```

When you first started writing R code, you might have solved the problem with copy-and-paste:

```{r, eval = FALSE}
df$a[df$a == -99] <- NA
df$b[df$b == -99] <- NA
df$c[df$c == -98] <- NA
df$d[df$d == -99] <- NA
df$e[df$e == -99] <- NA
df$f[df$g == -99] <- NA
```

One problem with copy-and-paste is that it's easy to make mistakes. Can you spot the two in the block above? These mistakes are inconsistencies that arose because we didn't have an authorative description of the desired action (replace $-99$ with `NA`). Duplicating an action makes bugs more likely and makes it harder to change code. For example, if the code for a missing value changes from $-99$ to 9999, you'd need to make the change in multiple places.

To prevent bugs and to make more flexible code, adopt the "do not repeat yourself", or DRY, principle. Popularised by the ["pragmatic programmers"](http://pragprog.com/about), Dave Thomas and Andy Hunt, this principle states: "every piece of knowledge must have a single, unambiguous, authoritative representation within a system". FP tools are valuable because they provide tools to reduce duplication.

We can start applying FP ideas by writing a function that fixes the missing values in a single vector:

```{r, eval = FALSE}
fix_missing <- function(x) {
  x[x == -99] <- NA
  x
}
df$a <- fix_missing(df$a)
df$b <- fix_missing(df$b)
df$c <- fix_missing(df$c)
df$d <- fix_missing(df$d)
df$e <- fix_missing(df$e)
df$f <- fix_missing(df$e)
```

This reduces the scope of possible mistakes, but it doesn't eliminate them: you can no longer accidentally type -98 instead of -99, but you can still mess up the name of variable. The next step is to remove this possible source of error by combining two functions. One function, `fix_missing()`, knows how to fix a single vector; the other, `lapply()`, knows how to do something to each column in a data frame.

`lapply()` takes three inputs: `x`, a list; `f`, a function; and `...`, other arguments to pass to `f()`. It applies the function to each element of the list and returns a new list. `lapply(x, f, ...)` is equivalent to the following for loop:

```{r, eval = FALSE}
out <- vector("list", length(x))
for (i in seq_along(x)) {
  out[[i]] <- f(x[[i]], ...)
}
```

The real `lapply()` is rather more complicated since it's implemented in C for efficiency, but the essence of the algorithm is the same. `lapply()` is called a __functional__, because it takes a function as an argument. Functionals are an important part of functional programming.

We can apply `lapply()` to this problem because data frames are lists. We just need a neat little trick to make sure we get back a data frame, not a list. Instead of assigning the results of `lapply()` to `df`, we'll assign them to `df[]`. R's usual rules ensure that we get a data frame, not a list. Putting these pieces together gives us:

```{r, eval = FALSE}
fix_missing <- function(x) {
  x[x == -99] <- NA
  x
}
df[] <- lapply(df, fix_missing)
```

This code has five advantages over copy and paste:

* It's more compact.

* If the code for a missing value changes, it only needs to be updated in
  one place.

* It works for any number of columns. There is no way to accidentally miss a
  column.

* There is no way to accidentally treat one column differently than another.

* It is easy to generalise this technique to a subset of columns:

    ```{r, eval = FALSE}
    df[1:5] <- lapply(df[1:5], fix_missing)
    ```

The key idea is function composition. Take two simple functions, one which does something to every column and one which fixes missing values, and combine them to fix missing values in every column. Writing simple functions that can be understood in isolation and then composed is a powerful technique.

What if different columns used different codes for missing values? 
```{r}
# Generate a sample dataset
set.seed(727)
df2 <- data.frame(lapply(c(-99, -999, -9999),
                         function(x) replicate(2, sample(c(1:10, x), 6, rep = TRUE))))
names(df2) <- letters[1:6]
df2
```

You might be tempted to copy-and-paste:

```{r}
fix_missing_99 <- function(x) {
  x[x == -99] <- NA
  x
}
fix_missing_999 <- function(x) {
  x[x == -999] <- NA
  x
}
fix_missing_9999 <- function(x) {
  x[x == -999] <- NA
  x
}
```

As before, it's easy to create bugs. Instead we could add another argument:

```{r}
fix_missing <- function(x, na.value) {
  x[x == na.value] <- NA
  x
}
```
```{r}
```{r}
library(dplyr)

letter.counts <- sort(unique(nchar(states)))
names <- paste(letter.counts, "chars", sep = "-")
lapply(setNames(letter.counts, names),
       function(x) {
         states[nchar(states) == x] %>%
           casefold(upper = x %% 2)
       })
```