---
title: "lab5"
author: "Andrew Do"
date: "June 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(stringr)
library(readr)
knitr::opts_chunk$set(echo = TRUE)
```

## Getting Help for Packages

You can see what an entire package has to offer with
```{r, eval = F}
help(package = "stringr")
```


## Counting number of vowels

Consider the following vector `letrs` which contains various letters:
```{r}
# Garbage string
letrs <- "gfngenodgsyctyviqyyigeinyngbkwjdjqkrrollhpxdkfldlytympmcgmjynihusebtq"
```

If you were to count the number of vowels in `letrs` you would get the following counts:

- a: 0
- e: 3
- i: 4
- o: 2
- u: 1

This exercise will guide you through how to arrive at this frequency table.

```{r}
# Create a character vector of vowels.  Give it an appropriate name.


# Use str_count to get the counts.  Assign this an appropriate name.


# Give each element of the vector a name using the names function.
# The result should look like:
# a e i o u
# 0 3 4 2 1
```

## Exploring new functions: str_pad
Write a short description of what each of the following calls of `str_pad` is doing.  The last three examples were found at the bottom of the function's documentation.

1. In this call of `str_pad`, __YOUR DESCRIPTION__.
```{r}
str_pad("a", width = 4, side = "left", pad = "x")
```

2. In this call of `str_pad`, __YOUR DESCRIPTION__.
```{r}
str_pad("a", width = 3, side = "left")
```

3. In this call of `str_pad`, __YOUR DESCRIPTION__.
```{r}
str_pad(c("a", "abc", "abcdef"), 10)
```

4. In this call of `str_pad`, __YOUR DESCRIPTION__.
```{r}
str_pad("a", c(5, 10, 20))
```

5. In this call of `str_pad`, __YOUR DESCRIPTION__.
```{r}
str_pad("a", 10, pad = c("-", "_", " "))
```

## Regular Expressions

We read in the data from the csv file and extract the names column
```{r}
flags <- read_csv("flags.csv")
countries <- flags$name
```

You may have noticed that in the `countries` vector, words in a country's name are separated by `-`.  Use `str_replace_all` to replace the hyphens with spaces.  For example `"Vatican-City"` should be replaced with `"Vatican City"`

```{r}
# Your code here
```

Use `str_detect` to see if there are any countries with the world "island" in them, ignoring case.  You may also want to look up how to use `regex` as a way to ignore case.

```{r}
# Your code here
```

Use `str_subset` to find all countries that begin with a capital `A`.
```{r}
# Your code here
```

Find all countries that end with `stan`
```{r}
# Your code here
```

Use `str_replace_all` to replace the `$`'s in `baws` with `s`'s.  Recall that the `$` is a metacharacter.
```{r}
baws <- "Bo$$"
```

Use a character class (denoted by square brackets) to find all countries beginning with a vowel and ending with a consonant.
```{r}
# Fill in
```




