---
title: Making Maps With R
author:
output: html_document
---



# Making Maps with R {#map-making-in-R} 

```{r, include = FALSE}
library(knitr)
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(dplyr)
library(stringr)
library(rvest)
library(readr)
opts_chunk$set(fig.width=10,
               fig.height=7,
               out.width = "600px",
               out.height = "420px",
               fig.path = "lecture_figs/making-maps-")
```

## New Packages

* ggmap
* mapdata

## Intro

For a long time, R has had a relatively simple mechanism, via the `maps` package, for making simple outlines
of maps and plotting lat-long points and paths on them.

More recently, with the advent of packages like `sp`, `rgdal`, and `rgeos`, R has been acquiring much of the
functionality of traditional GIS packages (like ArcGIS, etc).  This is an exciting development, but not
always easily accessible for the beginner, as it requires installation of specialized external libraries
(that may, on some platforms, not be straightforward) and considerable familiarity with GIS concepts.

More recently, a third approach to convenient mapping, using `ggmap` has been developed that allows the tiling of 
detailed base maps from Google Earth or Open Street Maps, upon which spatial data may be plotted.
Today, we are going to focus on mapping using base maps from R's tried and true `maps` package and also using the
`ggmap` package.  We won't cover the more advanced GIS-related topics nor using `rgdal`, or `sp` to plot
maps with different projections, etc.  Nor will cover the somewhat more simplified
approach to projections using the `mapproj` package.

As in our previous explorations in this course, when it comes to plotting, we are going to completely
skip over R's base graphics system and head directly to Hadley Wickham's `ggplot2` package.  Hadley has
included a few functions that make it relatively easy to interact with the data in R's `maps` package, and
of course, once a map layer is laid down, you have all the power of ggplot at your fingertips to overlay
whatever you may want to over the map.  `ggmap` is a package that goes out to different map servers and
grabs base maps to plot things on, then it sets up the coordinate system and writes it out as the base layer
for further ggplotting.  It is pretty sweet, but does not support different projections.

### Today's Goals

1. Introduce readers to the map outlines available in the `maps` package
    + Show how to convert those data into data frames that `ggplot2` can deal with
    + Discuss some `ggplot2` related issues about plotting things.
2. Use `ggmap` to make some pretty decent looking maps

I feel that the above twp topics should cover a large part of what people will need for making
useful maps of field sites, or sampling locations, or fishing track lines, etc. 

For today we will be skipping how to read in traditional GIS "shapefiles" so as to minimize
the number of packages that need installation, but keep in mind that it isn't too hard to do that
in R, too.

## Plotting maps-package maps with ggplot  {#maps-package-and-ggplot}

### The main players:

* The `maps` package contains a lot of outlines of continents, countries, states, and counties that have
been with R for a long time.  
* The `mapdata` package contains a few more, higher-resolution outlines.
* The `maps` package comes with a plotting function, but, we will opt to use `ggplot2` to plot the 
maps in the `maps` package.  
* Recall that `ggplot2` operates on data frames.  Therefore we need some way to
translate the `maps` data into a data frame format that `ggplot` can use.

### Maps in the maps package

* Package `maps` provides lots of different map outlines and points for cities, etc.  
* Some examples: `usa`, `nz`, `state`, `world`, etc.

### Making data frames from map outlines

* `ggplot2` provides the `map_data()` function.
    + Think of it as a function that turns a series of points along an outline into a data frame
    of those points.
    + Syntax:  `map_data("name")` where "name" is a quoted string of the name of a map in the `maps` or `mapdata`
    package
* Here we get a USA map from `maps`:
```{r}
usa <- map_data("usa")

dim(usa)

head(usa)

tail(usa)
```
* Here is the high-res world map centered on the Pacific Ocean from `mapdata`
```{r}
w2hr <- map_data("world2Hires")

dim(w2hr)

head(w2hr)

tail(w2hr)
```

### The structure of those data frames
These are pretty straightforward:

* `long` is longitude.  Things to the west of the prime meridian are negative.
* `lat` is latitude.
* `order`. This just shows in which order `ggplot` should "connect the dots"
* `region` and `subregion` tell what region or subregion a set of points surrounds.
* `group`.  This is _very important_!  `ggplot2`'s functions can take a group argument which 
controls (amongst other things) whether adjacent points should be connected by lines.  If they are
in the same group, then they get connected, but if they are in different groups then they don't.
    + Essentially, having to points in different groups means that `ggplot` "lifts the pen" when going between
    them.
    

### Plot the USA map

* Maps in this format can be plotted with the polygon geom.  i.e. using `geom_polygon()`.
* `geom_polygon()` drawn lines between points and "closes them up" (i.e. draws a line from the last
point back to the first point)
* You have to map the `group` aesthetic to the `group` column
* Of course, `x = long` and `y = lat` are the other aesthetics.

#### Simple black map
By default, `geom_polygon()` draws with no line color, but with a black fill:
```{r}
usa <- map_data("usa") # we already did this, but we can do it again
ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group)) + 
  coord_fixed(1.3)
```

#### What is this coord_fixed()?

* This is very important when drawing maps.
* It fixes the relationship between one unit in the $y$ direction and one unit in the $x$ direction.
* Then, even if you change the outer dimensions of the plot (i.e. by changing the window size or the size
of the pdf file you are saving it to (in `ggsave` for example)), the _aspect ratio_ remains unchanged.
* In the above case, I decided that if every $y$ unit was 1.3 times longer than an $x$ unit, then  the
plot came out looking good.
    + A different value might be needed closer to the poles.
    
#### Mess with line and fill colors

* Here is no fill, with a red line.  Remember, fixed value of aesthetics go _outside_ the `aes` function.
```{r}
ggplot(usa) + 
  geom_polygon(aes(x = long, y = lat, group = group), 
               fill = NA, color = "red") + 
  coord_fixed(1.3)
```
    
* Here is violet fill, with a blue line.
```{r}
gg1 <- ggplot(usa) + 
  geom_polygon(aes(x = long, y = lat, group = group), 
               fill = "violet", color = "blue") + 
  coord_fixed(1.3)
gg1
```
    
#### Adding points to the map

* Let's add black and yellow points at our lab and at the NWFSC in Seattle.
```{r}
labs <- data.frame(
  long = c(-122.064873, -122.306417),
  lat = c(36.951968, 47.644855),
  names = c("SWFSC-FED", "NWFSC"),
  stringsAsFactors = FALSE
  )  

gg1 + 
  geom_point(data = labs, aes(x = long, y = lat), color = "black", size = 5) +
  geom_point(data = labs, aes(x = long, y = lat), color = "yellow", size = 4)
```

#### See how important the group aesthetic is

To show how important the group aesthetic is, here we plot the map without using the group aesthetic:
```{r}
ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat), fill = "violet", color = "blue") + 
  geom_point(data = labs, aes(x = long, y = lat), color = "black", size = 5) +
  geom_point(data = labs, aes(x = long, y = lat), color = "yellow", size = 4) +
  coord_fixed(1.3)
```

### State maps
We can also get a data frame of polygons that tell us above state boundaries:
```{r}
states <- map_data("state")
dim(states)
head(states)
tail(states)
```

#### Plot all the states, all colored a little differently

This is just like it is above, but we can map fill to `region` and 
make sure the the lines of state borders are white.
```{r}
ggplot(states) + 
  geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + 
  coord_fixed(1.3) +
  guides(fill = FALSE)  # do this to leave off the color legend
```

Boom! That is easy.

#### Plot just a subset of states in the contiguous 48:

* We can use it to grab just CA, OR, and WA:

```{r}
west_coast <- filter(states, region %in% c("california", "oregon", "washington"))

ggplot(west_coast) + 
  geom_polygon(aes(x = long, y = lat), fill = "palegreen", color = "black") 
```

#### Man that is ugly!! 

* I am just keeping people on their toes. What have we forgotten here?
    + `group`
    + `coord_fixed()`
* Let's put those back in there:
```{r}
ggplot(data = west_coast) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill = "palegreen", color = "black") + 
  coord_fixed(1.3)
```

Phew! That is a little better!


#### Zoom in on California and look at counties

* Getting the california data is easy:
```{r}
ca_df <- filter(states, region == "california")

head(ca_df)
```

* Now, let's also get the county lines there
```{r}
counties <- map_data("county")
ca_county <- filter(counties, region == "california")

head(ca_county)
```

* Plot the state first but let's ditch the axes gridlines, and gray background by
using the super-wonderful `theme_nothing()`.

```{r}
ca_base <- ggplot(ca_df, aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")
ca_base + theme_nothing()
```

* Now plot the county boundaries in white:
```{r}
ca_base + theme_nothing() + 
  geom_polygon(data = ca_county, fill = NA, color = "white") +
  geom_polygon(color = "black", fill = NA)  # get the state border back on top
```

#### Get some facts about the counties

* The above is pretty cool, but it seems like it would be a lot cooler if we could plot some information about
those counties.  
* Now I can go to [wikipedia](http://www.california-demographics.com/counties_by_population) and grab population and area data for each county.

```{r, warning=FALSE, message=FALSE}
    
# scrape population data
county_html <- read_html("https://en.wikipedia.org/wiki/List_of_counties_in_California")
pop_and_area <- county_html %>% 
  html_node(".wikitable.sortable") %>% 
  html_table() %>% 
  select(county = County, population = `Population[6]`, area = `Area[4]`) %>% 
  mutate(county = county %>% 
           str_replace_all(c("City and County of San Francisco" = "San Francisco",
                         " County" = "")) %>% 
           str_to_lower,
         population = population %>% 
           str_replace_all(".*\u2660", "") %>% parse_number,
         area = area %>% 
           str_replace_all(".*\u2660|\n.*", "") %>%  parse_number)
```
* We now have the numbers that we want, but we need to attach those to 
every point on polygons of the counties.  This is a job for `inner_join`.

```{r}
cacopa <- inner_join(ca_county, pop_and_area, by = c("subregion" = "county"))
```
* And finally, add a column of `people_per_mile`:
```{r}
cacopa <- mutate(cacopa, people_per_mile = population / area)

head(cacopa)
```

#### Now plot population density by county

If you were needing a little more elbow room in the great Golden State, this shows you where you can find it:
```{r}
# prepare to drop the axes and ticks but leave the guides and legends
# We can't just throw down a theme_nothing()!
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()
  )

elbow_room1 <- ca_base + 
      geom_polygon(data = cacopa, aes(fill = people_per_mile), color = "white") +
      geom_polygon(color = "black", fill = NA) +
      theme_bw() +
      ditch_the_axes

elbow_room1 
```

#### Lame!

* The popuation density in San Francisco is so great that it makes it hard to discern differences between
other areas.
* This is a job for a scale transformation.  Let's take the log-base-10 of the population density.
* Instead of making a new column which is log10 of the `people_per_mile` we can just apply the
transformation in the gradient using the `trans` argument
```{r}
elbow_room1 + scale_fill_gradient(trans = "log10")
```

 
#### Still not great
It's very common for color to represent a dimension of your data in spatial visualizations.  As such, you should think about the color palette carefully.  Here, contrasting colors are used to show the gamut of population density in California.
```{r}
eb2 <- elbow_room1 + 
    scale_fill_gradient(low = "#0C374D", high = "#EFD469", trans = "log10")
eb2
```


### Zooming in?
Note that the scale of these maps from package `maps` are not great. We can zoom in to the 
Bay region, and it sort of works scale-wise, but if we wanted to zoom in more, it would
be tough.  

Let's try!
```{r}
eb2 + xlim(-123, -121.0) + ylim(36, 38)
```

* Welp, that was a big dose of derpatitis. Why?
* Recall that `geom_polygon()` connects the end point of a `group` to its starting point.
* `xlim` and `ylim` functions in `ggplot2` discard all the data that is not within the plot area.  
    + Hence there are new starting points and ending points for some groups (or in this case the
    black-line permiter of California) and those points get connected.  Not good.


### True zoom.

* If you want to keep all the data the same but just zoom in, you can use the `xlim` and `ylim` arguments to `coord_cartesian()`.  Though, to keep the aspect ratio correct we must use `coord_fixed()` instead of 
`coord_cartesian()`.
* This chops stuff off but doesn't discard it from the data set:
```{r}
eb2 + coord_fixed(xlim = c(-123, -121.0),  
                  ylim = c(36, 38), 
                  ratio = 1.3)
```


## ggmap

The `ggmap` package is the most exciting R mapping tool in a long time.  You might be able to get better looking maps at some resolutions by using shapefiles and rasters from naturalearthdata.com but `ggmap` will get you 95% of the way there with only 5% of the work!

### Three examples

1. Named "sampling" points on the Sisquoc River from the "Sisquoctober Adventure"
2. A GPS track from a short bike ride in Wilder Ranch.
3. Fish sampling locations from the coded wire tag data base.
    
### How ggmap works

* ggmap simplifies the process of downloading base maps from Google or Open Street Maps or Stamen Maps
to use in the background of your plots.
* It also sets the axis scales, etc, in a nice way.
* Once you have gotten your maps, you make a call with `ggmap()` much as you would with `ggplot()`
* Let's do by example.

### Sisquoctober

* Here is a small data frame of points from the Sisquoc River.
```{r}
sisquoc <- read_tsv("data/sisquoc-points.txt")
sisquoc
```

Let's try using the zoom level.  Zoom levels go from 3 (world scale to 20 (house scale)).
```{r}
# compute the mean lat and lon
ll_means <- summarise_each(sisquoc, "mean", lon, lat) %>% as.numeric(.[1, ])
sq_map2 <- get_map(location = ll_means,  maptype = "satellite", 
                   source = "google", zoom = 15)

ggmap(sq_map2) + 
  geom_point(data = sisquoc, color = "red", size = 4) +
  geom_text(data = sisquoc, aes(label = name), 
            angle = 60, hjust = -.25, color = "yellow")
```
* That is decent.  How about if we use the "terrain" type of map:
```{r}
sq_map3 <- get_map(location = ll_means,  maptype = "terrain", 
                   source = "google", zoom = 15)
ggmap(sq_map3) + 
  geom_point(data = sisquoc, color = "red", size = 4) +
  geom_text(data = sisquoc, aes(label = name), 
            angle = 60, hjust = -.25, color = "yellow")
```
    
* That is cool, but I would search for a better color for the lettering...

### How about a bike ride?

* Here we use some GPS data for a bike ride.
* We can plot the route like this:
```{r}
bike <- read_csv("data/bike-ride.csv")
head(bike)

bikemap1 <- get_map(location = c(-122.080954, 36.971709), 
                    maptype = "terrain", source = "google", zoom = 14)
ggmap(bikemap1) + 
  geom_path(data = bike, aes(color = elevation), size = 3, lineend = "round")
```

* Note that getting the right zoom and position for the map is sort of trial and error.  You can go to google maps to figure out where the center should be (right click and choose "What's here?" to get the lat-long of any point. )

### Fish sampling locations

In this example, we use fishery data from Eric Anderson of NOAA.  He whittled down some stuff in the coded wire tag data base to georeferenced marine locations in British Columbia where at least one Chinook salmon was recovered in between 2000 and 2012 inclusive.  To see how he did all that you can check out [this](https://github.com/eriqande/pbt-feasibility/blob/4ea2fc960f74f66b5ec3a11c107cdc52bfb346dc/Rmd/02-02-explore-recovery-and-catch-sample-data.Rmd#looking-at-locations-of-location-codes)

Let's have a look at the data:
```{r}
bc <- readRDS("data/bc_sites.rds")

# look at some of it:
bc %>% select(state_or_province:sub_location, longitude, latitude)
```

So, we have 1,113 points to play with.  

#### What do we hope to learn?

* These locations in BC are hierarchically structured.  I am basically interested in how close together sites in the same "region" or "area" or "sector" are, and pondering whether it is OK to aggregate fish recoveries at a certain level for the purposes of getting a better overall estimate of the proportion of fish from different hatcheries in these areas.
* So, pretty simple stuff.  I just want to plot these points on a map, and paint them a different
color according to their sector, region, area, etc.
* Let's just enumerate things first:

```{r}
bc %>% group_by(sector, region, area) %>% tally()
```

* That looks good.  It appears like we could probably color code over the whole area down to region, and
then down to area within subregions.

#### Making a map.

* Let's try to use `make_bbox()`.  This makes a bounding box rather than relying on the zoom feature.  It's an experimental feature that would be cool if it worked reliably, but unfortunately, it's known to fail.

```{r}
# compute the bounding box
bc_bbox <- make_bbox(lat = latitude, lon = longitude, data = bc)
bc_bbox

# grab the maps from google
bc_big <- get_map(location = bc_bbox, source = "google", maptype = "terrain")

# plot the points and color them by sector
ggmap(bc_big) + 
  geom_point(data = bc, 
             mapping = aes(x = longitude, y = latitude, color = sector))
```

* Cool! That was about as easy as could be.  North is in the north, south is in the south, and 
the three reddish points are clearly aberrant ones at the mouths of rivers.


#### Coloring it by region

```{r}
ggmap(bc_big) + 
  geom_point(data = bc, mapping = aes(x = longitude, y = latitude, color = region))
```

#### Zooming in on each region and coloring by area

* It is time to really put this thing through its paces.  (Keeping in mind that `make_bbox()` might fail...)
* I want to make series of maps.  One for each region, in which the the areas
in that region are colored differently.
* How?  Let's make a function:  you pass it the region and it makes the plot.

```{r}
region_plot <- function(my_region) {
  region_data <- bc %>% filter(region == my_region)
  bbox <- make_bbox(lon = longitude, lat = latitude, data = region_data)
  my_map <- get_map(location = bbox, source = "google", maptype = "terrain")
  # now we want to count up how many areas there are
  num_areas <- region_data %>% summarise(n_distinct(area)) %>% as.numeric
  num_points <- nrow(region_data)
  
  the_map <- ggmap(my_map) +
    geom_point(data = region_data, 
               mapping = aes(x = longitude, y = latitude), 
               size = 4, color = "black") +
    geom_point(data = region_data, 
               mapping = aes(x = longitude, 
                             y = latitude, color = area), size = 3) +
    ggtitle(sprintf("BC Region: %s with %d locations in %d area(s)",
                    my_region, num_points, num_areas))
  
  ggsave(sprintf("bc_region%s.pdf", my_region), the_map, width = 9, height = 9)
}
```

So, with that function we just need to cycle over the regions and make all those plots.
Note that I am saving them to PDFs because it is no fun to make a web page with all of those in there.

```{r, eval = FALSE}
# invisible suppresses lapply's list output, which we're not interested in
bc$region %>% 
  unique %>% 
  lapply(region_plot) %>%
  invisible
```

## Further Reading

[ggmap cheat sheet](https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/ggmap/ggmapCheatsheet.pdf)
[ggmap documentation](https://cran.r-project.org/web/packages/ggmap/ggmap.pdf)
[Spatial Plots](https://journal.r-project.org/archive/2013-1/kahle-wickham.pdf)

