---
title       : Lesson 12
subtitle    : Factors and Graphing
author      : Andrew Do
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
```

## Factors

* A factor is a vector that can contain only predefined values, and is used to store categorical data.  They handle categorical data.
* Examples Include:
  - Treatment A, Treatment B, Treatment C
  - Small, Medium, Large
  - Likert Scale (Very bad, bad, somewhat bad, neutral, ...)

---

## Internals of Factors

* Factors are NOT basic data types
* Internally, they are stored as integers

```{r}
my_factor <- factor(c("A", "A", "B", "A", "C"))

typeof(my_factor)
levels(my_factor)
```

---

## More Internals of Factors

Levels keep track of what the integers actually mean

```{r}
str(my_factor)
levels(my_factor)
```

---

## 3 types of categorical variables

### Binary
Two categories (e.g. Homeowner or not).  Usually handled by logical vectors

### Nomimal
No ordering to categories (e.g. Treatment A, B, C).  Handled by factors

### Ordinal
Natural ordering to categories (slow, average, fast).  Handled by ordered factors

---

## Creating Factors

Create a factor using `factor()`
```{r}
factor(c("red", "blue", "red", "red", "yellow"))
```

The values a factor can take are called __levels__.  It's not necessary for all levels to be present in a factor vector.
```{r}
factor(c("red", "blue", "red", "red", "yellow"), 
       levels = c("red", "yellow", "blue", "green"))
```

---

## Ordered factors

To create an ordered factor, we just need to specify the levels in order and then pass `TRUE` to the `ordered` argument.

```{r}
x <- factor(c("sm", "lg", "lg", "md", "sm"), levels = c("sm", "md", "lg"), ordered = TRUE)
x 
```

This allows us to sort the factors
```{r}
sort(x)
```

---

## More Internals of Factors

To see what integers represent each level, use `unclass()`.  The order of the attributes gives the integer values.
```{r}
x <- factor(c("sm", "lg", "lg", "md", "sm"), levels = c("sm", "md", "lg"), ordered = TRUE)
unclass(x)
```

To change the levels, you can assign new values a `levels()` call:
```{r}
levels(x) <- c("S", "M", "L")
x
```

---

## stringsAsFactors

* By default, `data.frame` converts all strings to factors
* This will often mess things up (as we saw in lab)
* To fix this, specify `stringsAsFactors = FALSE` when making data frames and then go back and turn specific columns into factors as needed.
* By default, `readr` functions such as `read_csv` and `read_tsv` take care of this for you

```{r}
df <- data.frame(x = c("a", "b", "c"), y = 1:6)
str(df)
```

---

## Data Set

For this next exercise, consider this data frame
```{r}
tbl_df(mtcars)
```

---

## How would you make this graph?
Talk with the people around you.

```{r, echo = FALSE, fig.width=10, fig.height=5.5}
ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point(aes(color = factor(cyl)), size = 3) +
  geom_smooth() +
  ggtitle("Horsepower and Fuel Efficiency") +
  theme(title = element_text(size = 26),
        axis.text = element_text(size = 18),
        axis.title = element_text(size = 22),
        legend.text = element_text(size = 18))
```

---

## How would you make this graph?

```{r, echo = F, fig.width=7.5, fig.height=5.5}
mtcars %>%
  mutate(cyl = factor(cyl)) %>%
  ggplot(aes(x = cyl, y = mpg, fill = cyl)) +
    geom_boxplot() +
    ggtitle("Horsepower and Fuel Efficiency") +
    theme(title = element_text(size = 26),
          axis.text = element_text(size = 18),
          axis.title = element_text(size = 22),
          legend.text = element_text(size = 18))
```

---

## Example Code

### We will want to formalize how we talk about graphs
```{r, eval = FALSE}
mtcars %>%
  mutate(cyl = factor(cyl)) %>%
  ggplot(aes(x = cyl, y = mpg, fill = cyl)) +
    geom_boxplot() +
    ggtitle("Horsepower and Fuel Efficiency") +
    theme(title = element_text(size = 26),
          axis.text = element_text(size = 18),
          axis.title = element_text(size = 22),
          legend.text = element_text(size = 18))
```
 
---

## TOMMOROW

* Formalizing the grammar of graphics
* Basic ggplot2
* To do tonight: Install the ggplot2 package