---
title       : Lesson 5
subtitle    : Regular Expressions R
author      : Andrew Do
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Motivation

```{r libraries, echo = F}
library(stringr)
```

* How would you describe a (US) phone number to someone?
* How about the date format?
* Time?
* Money?
* Email address?

---

## Pattern Matching

* Each of the data I just asked about has a fairly _regular_ pattern that we can pick out by eye

* In this lesson, we're going to learn about __regular expressions__, a precise and concise language for describing patterns.

---

## Phone numbers

With a partner, spend 5 minutes discussing in detail:

* How are phone numbers normally written?
* How can we describe this format?

Write down this description in words.

---

## Phone numbers

* 10 digits, grouped 3-3-4
* Separated by a space, short dash, or parentheses

* How do we translate that into something the computer can understand? Regular expressions:

  * `[0-9]` matches any number between 0 and 9
  * `[- ()]` matches `-`, `SPACE`, `(`, or `)`
  * `{3}` following a character means that that character is repeated 3 times.

---

## Test text

```{r}
email <- "Nothing. It is a generic folder that has Enron Global Markets on the cover. It is the one that I sent you to match your insert to when you were designing. With the dots. I am on my way over for a meeting, I'll bring one. 

Juli Salvagio
Manager, Marketing Communications
Public Relations
Enron-3641
1400 Smith Street
Houston, TX 77002
713-345-2908-Phone
713-542-0103-Mobile
713-646-5800-Fax"
```

---

```{r}
phone <- "[ (][0-9][0-9][0-9][- )][0-9][0-9][0-9][- ]
[0-9][0-9][0-9][0-9]"
phone2 <- "[0-9]{3}[- .][0-9]{3}[- .][0-9]{4}"

str_detect(email, phone2)
str_extract(email, phone2)
str_extract_all(email, phone2)
```

---

# Qualifiers
The following special symbols, called __metacharacters__ have special meanings in regular expressions.

Qualifier | >=   | <=   | description
--------- | ---- | ---- | ----------------------
?         | 0    | 1    | at most 1
+         | 1    | Inf  | at least 1
*         | 0    | Inf  | any number of
{m,n}     | m    | n    | between m and n
{,n}      | 0    | n    | at most n
{m,}      | m    | Inf  | at least m
{n}       | n    | n    | exactly n

---

# Other metacharacters
These symbols also represent something other than their _literal_ meaning:

Metacharacter | Meaning   | Example Usage | Description
------------- | --------- | ------------- | -------------------
()            | group     | "(ham){3}"    | matches "hamhamham"
.             | anything  | ".{3}"        | matches any 3 characters (e.g "ham", "h@m", "h49")
^             | beginning | "\^ham"        | only matches "ham" if it appears at the beginning of the string
$             | end       | "ham$"        | only matches "ham" if it appears at the end of the string

---

# Escaping

* R uses the double-slash `\\` to __escape__ special meanings
* Illustrating Examples
  
  * "a.b" matches the letter a, then any character, then the letter b
  * "a\\\\.b" matches the literal string "a.b"

---

# Character Classes

* Square brackets also have a special meaning in regular expressions
* They denote something called a __character class__
* You've already seen two examples: `[0-9]` matches any one digit. `[( -]` matches `(`, `-`, or `-`
* `[a-z]` matches any one lowercase letter
* `[a-zA-Z]` matches any one letter
* `[LATE]` matches one of the four letters `L`, `A`, `T`, `E`.
* `[LATE]{4}` matches `"LATE"`, but it also matches `"TEAL"`
```{r}
char_class <- "[LATE]{4}"
str_detect("TEALING", char_class)
```

---

# Character classes

* All of the metacharacters mentioned lose their special meanings inside of character classes
* The only exception is `^`, which when placed at the beginning means "everything but the characters in this class"
```{r}
# Everything but the letters E, A, and T
neg_class <- "[^EAT]"

# BAT contains B, which is not E, A, or T
str_detect("BAT", neg_class)

# Beat does not contain those characters either
str_detect("Beat", neg_class)
```

---

# Your Turn

### Create a regular expression to match a date. Test it against the following cases:
```{r, eval = F}
c("10/14/1979", "20/20/1945", "1/1/1905", "5/5/5")
```

### Create a regular expression to match a time. Test it against the following cases:
```{r, eval = F}
c("12:30 pm", "2:15 AM", "312:23 pm", "1:00 american", "08:20 am")
```

