---
title: "Creating a Shiny App Part 2"
author: "Andrew Do"
date: "July 14, 2016"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Recap
Yesterday, you created an interface that looked like:

```{r, eval = FALSE}
shinyUI(fluidPage(
  titlePanel("Occupancy Rates in Berkeley"),
  
  sidebarLayout(
    sidebarPanel(
      helpText("Upload an XML file with the Census Data to get started"),
      
      fileInput("file", label = "File input"),
      
      selectInput("neigh", 
        label = "Choose the neighborhood to inspect",
        choices = list("North Berkeley", 
                       "Rockridge",
                       "Downtown Berkeley", 
                       "Berkeley Hills"),
        selected = "North Berkeley"),
      
      sliderInput("occupancy", 
        label = "Number of People in Household:",
        min = 0, max = 40, value = c(0, 40))
    ),
    
    mainPanel()
  )
))
```

Our goal today is to learn how to make the interface actually do something.  Start by copying the app you submitted yesterday into today's folder.

## Creating a Reactive Object

We can break the process of creating reactive output into two steps:

1. Add an R object to yoru user-interface with `ui.R`
2. Tell Shiny how to build the object in `server.R`.  The object will be reactive if the code that builds it calls a widget value.

### Step 1: Add an R object to the UI.  

To add an object to the UI, you simply call on one of the functions below, placing it in the Panel that you want it to appear in.

Function                | Type of Output
----------------------- | -----------------------------------
htmlOutput              | raw html
imageOutput             | image
plotOutput              | plot
tableOutput             | table
textOutput              | text
uiOutput                | raw HTML
verbatimTextOutput      | unformatted text

Each of these functions have the same required first argument: a character string that tells the UI what object from the `server.R` script it should be looking for.  In the example below, `textOutput` will be looking for the contents of `output$text1`.

```{r, eval = FALSE}
shinyUI(fluidPage(
  titlePanel("Occupancy Rates in Berkeley"),
  
  sidebarLayout(
    sidebarPanel(
      helpText("Upload an XML file with the Census Data to get started"),
      
      fileInput("file", label = "File input"),
      
      selectInput("neigh", 
        label = "Choose the neighborhood to inspect",
        choices = list("North Berkeley", 
                       "Rockridge",
                       "Downtown Berkeley", 
                       "Berkeley Hills"),
        selected = "North Berkeley"),
      
      sliderInput("occupancy", 
        label = "Number of People in Household:",
        min = 0, max = 40, value = c(0, 40))
    ),
    
    mainPanel(
      textOutput("text1")
    )
  )
))
```

### Creating an Object in the Server Script

Now that you've told the UI where to display the object, you'll actually need to build it server-side.

We modify the `server.R` code from yesterday a little.

```{r, eval = FALSE}
shinyServer(function(input, output) {
  
  # Note the curly braces in addition to the parentheses
  output$text1 <- renderText({ 
    "You have selected this"
  })
  
})
```

The `function(input, output)` call plays a very special role in the `Shiny` app.  It will build an `output` object with named elements that can be accessed by the UI.  Notice that we assign the text to `output$text1` in the server script when we want it to output to `"text1"` on the UI.

The other thing that you've probably noticed is that I had to wrap the expression in `renderText`.  This does some light pre-processing that's not super-important to understand.  What's important is to remember that every `*Output` function in the UI needs a corresponding `render*` on the server.  You should call them with parentheses and curly braces to group the expression.

Function            | Object Type
------------------- | -------------------------------------------------
renderImage         | images (usually as a path on your computer)
renderPlot          | plots
renderPrint         | any printed output
renderTable         | tabular structures like data frames and matrices
renderText          | character strings
renderUI            | a Shiny tag object or HTML

### Using Widget Values Notice 

Just like how the `ui.R` script can take `output` values from the `server.R` script, the `server.R` process can access `input` values given by widgets in the UI.  In our example, we have access to three `input` values: `input$file`, `input$neigh`, and `input$occupancy`, each derived from their corresponding widgets.  We can modify the server script as follows to take the widget values into account:

```{r, echo = FALSE}
ui <- shinyUI(fluidPage(
  titlePanel("Occupancy Rates in Berkeley"),
  
  sidebarLayout(
    sidebarPanel(
      helpText("Upload an XML file with the Census Data to get started"),
      
      fileInput("file", label = "File input"),
      
      selectInput("neigh", 
        label = "Choose the neighborhood to inspect",
        choices = list("North Berkeley", 
                       "Rockridge",
                       "Downtown Berkeley", 
                       "Berkeley Hills"),
        selected = "North Berkeley"),
      
      sliderInput("occupancy", 
        label = "Number of People in Household:",
        min = 0, max = 40, value = c(0, 40))
    ),
    
    mainPanel(
      textOutput("text1"),
      textOutput("text2")
    )
  )
))


server <- shinyServer(function(input, output) {
  
  # Note the curly braces in addition to the parentheses
  output$text1 <- renderText({ 
    paste("You have selected", input$neigh, "as your neighborhood.", sep = " ")
  })
  output$text2 <- renderText({ 
    paste("You are interested in houses where",
          input$occupancy[1], "to", input$occupancy[2],
          "people live.", sep = " ")
  })
  
})

shinyApp(ui, server)
```

### String Evaluations

This is a very cursory introduction into a deep topic called "non-standard evaluation".  For more information, run
```{r, eval = FALSE}
vignette("nse")
```


`dplyr`, `tidyr`, and `ggplot2` allow you to call on variables hidden inside data frames without having to do something like `df$var1`.  This is great for interactive work when you're the one providing names to the function, but what happens when you want reactive content---when you want `R` to adjust the function calls automatically based on user input? In comes string evaluation.

Every `dplyr` and `tidyr` `verb()` function comes with its cousin, a `verb_()` string-evaluated function.

```{r}
library(dplyr)
df <- data.frame(FirstVar = 1:3, SecondVar = 4:6)

# These two are equivalent
select(df, FirstVar)
select_(df, .dots = list("FirstVar"))

# When the second is more useful
variable_of_interest <- "FirstVar"
# select(df, variable_of_interest) # Does not work
select_(df, .dots = list(variable_of_interest)) # This does work
```

Very similarly, `ggplot2` has an equivalent `aes_string` as a string-evaluated version of `aes`.

```{r}
library(ggplot2)

var1 <- "FirstVar"
var2 <- "SecondVar"

# This doesn't work
# ggplot(df) +
#   geom_point(aes(x = var1, y = var2))

# This does
ggplot(df) +
  geom_point(aes_string(x = var1, y = var2))
```

### Lab Work
Using the `diamonds` dataset (included when you load `ggplot2`), write a Shiny App that plots a scatter plot of `price` against either depth, table, or carat as per the user's choice via a dropdown menu.  The user should be able to select `cut` and `color` via checkboxes to subset the data before plotting.  For example, if they choose `carat`, Ideal `cut`, and E and I `color`, then they should see the following plot:

```{r}
diamonds %>% 
  # The cut and color values will vary with user choice
  filter(cut == "Ideal", color %in% c("E", "I")) %>% 
  # The x aesthetic should be dynamic
  ggplot(aes(x = carat, y = price)) +
  geom_point()
```

On the UI side, you'll use `outputPlot`.  Consequently, you'll wrap your `ggplot` call in `renderPlot` server-side.  When you're done, submit your `server.R` and `ui.R` files.
