---
title: "Midterm Topics"
author: "Andrew Do"
date: "June 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exam Logistics

* The exam will take place on Thursday, June 30 during our usual classtime in our usual classroom.
* The exam should take most people 80 minutes, but you will be allowed 120 minutes.
* You will take the exam in ink, crossing out unwanted answers with a single strike-through
* You are allowed a single cheat sheet.  It can be in the form of flashcards, 8.5x11 paper, mobius strips---so long as the surface area covered in notes including the whitespace bounding box is less than $93.5 in^2$ (8.5x11 piece of paper).  The scribing method doesn't matter---it can be typed, handwritten, printed in 5 overlayed ink colors decipherable only with color-filtering glasses.

## Basic Terminal

1.  You should recognize basic terminal commands and be able to describe what they do
2.  You should be able to construct absolute and relative paths given a directory tree

## Data Types

1.  You should be able to name the basic data types in `R` and give examples of each: numeric, integer, complex, logical, character
2.  You should know the difference between atomic (e.g. vectors and matrices) and non-atomic classes (e.g. lists and data frames)
3.  You should know basic interactions with factors and time data.
4.  You should be able to distinguish between an instant, period, duration, and interval.

## Reading documention

You might be asked to read documentation on a function that you've never seen before.  You could then be asked to write a basic function call that would return a requested result.  This will require understanding data types.

## Subsetting

You should be able to subset vectors, lists, matrices, and data frames by index and by name.  Know which data type each subsetting operation returns.  Below is a nonexhaustive list of examples.

Subset Call           | Meaning
--------------------- | ---------------------------------------------------
myvector[1]           | vector containing first element of myvector
myvector[c(3,6)]      | vector containing third and sixth element of myvector
myvector["bob"]       | element named "bob" in myvector
mylist[2]             | list containing second element of mylist
mylist[[2]]           | second element of mylist
mylist[2:3]           | list containing second and third elements of mylist
mymatrix[1,2]         | element in first row, second column of mymatrix
mymatrix[1, ]         | first row of mymatrix
mymatrix[ ,3]         | third column of mymatrix
mydataframe["x"]      | data frame containing "x" column of mydataframe
mydataframe[["x"]]    | vector containing data from the "x" column of mydataframe
mydataframe$x         | vector containing data from the "x" column of mydataframe

## Strings

1.  You should be able to construct regular expressions
2.  Save space on your cheat sheet: memorize the metacharacters---you should be super comfortable with these anyway.
3.  You don't have to know the specific arguments of each of the `stringr` functions, but you should remember what each one does

Function       | Description
-------------- | -------------------------------
str_c          | same usage as paste
str_length     | number of characters
str_sub        | extracts substrings
str_dup        | duplicates characters
str_to_lower   | converts to lowercase
str_to_upper   | converts to uppercase
str_pad        | pads a string
str_detect     | detects presence of a pattern
str_locate     | finds first position of pattern
str_replace    | replaces first matched pattern
str_replace_all| replace all matched patterns
str_match      | returns first matched pattern
str_match_all  | returns all matched patterns
str_split      | splits string at matched pattern

## Control Flow and user-defined functions

1.  You should be able to construct for and while loops
2.  You should understand how to use if, else, ifelse
3.  You should be able to read through a basic function and predict output based off a function call
4.  You should understand how `lapply`, `sapply`, `vapply`, and `tapply` work.

```{r, eval = F}
x <- -1
y <- 2

example <- function(x = 0) {
  y <- x + y
  x + y
}
# What does example() return? example(1)? example(x = y)?
```


## Data Manipulation

1.  You should be very comfortable with the data verbs

Verb        | Meaning
----------- | -----------------------------------------
filter      | keep rows matching criteria
select      | pick columns by name
arrange     | reorder rows
mutate      | create new variables/modify existing ones
summarise   | reduce variables to values
gather      | Take columns and collapse into key-value pairs
spread      | Take key-value pair and lay it over multiple columns
separate    | Turn a single character column into multiple columns
unite       | Turn multiple character columns into one

2.  You should understand how joins work

Type      | Description
--------- | ------------------------------------------------------------------------------
inner     | return rows in `x` AND `y`
left      | return all of `x` and matching rows in `y`
semi_join | return rows of `x` that match `y`, only columns of `x`
anti_join | only rows of `x` not in `y`, keeping only columns of `x`
full_join | return all rows and all columns from both `x` and `y`, matching when possible

## Grammar of Graphics

1.  Be able to distinguish between variable-mapped vs. fixed aesthetics in a graph.
2.  Be able to identify glyphs and their variable mappings
3.  Be able to identify parts of a graph
4.  Know how faceting works