---
title: "Lesson 1"
subtitle: "A Brief Introduction to R"
output:
  tufte::tufte_html: default
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

# Introduction

Here are some very basic things you can do in R:

R can be used as a calculator: any expressions that you type into the R console will be
computed, and the answer printed.  For example

```{r}
1 + 1
5 / 9
```

You can write comments in R using the hash symbol #.

```{r}
# This line of code will not get evaluated.  2 + 2
## The line below this one will be though
2 + 2
```

You can form vectors in R using the command `c`.  Operations are usually applied element-by-element to a vector.

```{r}
# Adding two vectors together
c(1, 2, 3) + c(4, 5, 6)

# Multiplying two vectors together
c(1, 2, 3) * c(1, 2, 3)
```

You can "save" objects in memory by using either the arrow `<-` or an equal sign `=`. We call this action "assignment".

```{r}
# Assigning the vector (1, 2, 3) to x
x <- c(1, 2, 3)
# Assigning the vector (4, 5, 6) to y
y <- c(4, 5, 6)
# Adding x and y should return (5, 7, 9)
x + y
```

All functions in R are called using their names followed by their arguments in parentheses (more on this later).  For example, here we take the average of `x`, which was assigned the values 1, 2, and 3 earlier.

```{r}
mean(x)
```

