---
title: "Lesson 1"
subtitle: "A Brief Introduction to R"
output:
  tufte::tufte_html: default
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

# Introduction

Here are some very basic things you can do in R:

## Calculator

R can be used as a calculator: any expressions that you type into the R console will be
computed, and the answer printed.  For example

```{r}
1 + 1
5 / 9
```

## Comments

You can write comments in R using the hash symbol #.

```{r}
# This line of code will not get evaluated.  2 + 2
## The line below this one will be though
2 + 2
```

## Vectors

You can form vectors in R using the command `c`.  Operations are usually applied element-by-element to a vector.

```{r}
# Adding two vectors together
c(1, 2, 3) + c(4, 5, 6)

# Multiplying two vectors together
c(1, 2, 3) * c(1, 2, 3)
```

## Naming Objects

You can "save" objects in memory by using either the arrow `<-` or an equal sign `=`. We call this action "assignment to a name".

```{r}
# Assigning the vector (1, 2, 3) to x
x <- c(1, 2, 3)
# Assigning the vector (4, 5, 6) to y
y <- c(4, 5, 6)
# Adding x and y should return (5, 7, 9)
x + y
```
There are some rules for naming objects in R:

* Names cannot start with a number, so, for example `001graph` would not be valid.
* Names cannot contain punctuation, so `n*mmatrix` is not legitimate.  The only two exceptions are periods `.` and underscores `_`.
* Names are case-sensitive, so `vorpal1` and `Vorpal1` are two different objects

## Functions

All functions in R are called using their names followed by their arguments in parentheses (more on this later).  For example, here we take the average of `x`, which was assigned the values 1, 2, and 3 earlier.

```{r}
mean(x)
```

# Getting Help in R

To get help about a function in R, you can use `help(function_name)` or `?function_name`.  For example, the following calls would fetch information about the mean function.

```{r, eval = F}
help(mean)
?mean
```

More often than not, though, I end up deferring to StackExchange or Google.