---
title: "lab9"
author: "Andrew Do"
date: "June 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

The following libraries are allowed for this lab.  They have already been loaded in this document.

* `stringr`
* `dplyr`
* `tidyr`

```{r, include = FALSE}
library(stringr)
library(dplyr)
library(tidyr)
```


## Our Glorious Foundation

The vector pioneers has already been created for you
```{r}
pioneers <- c("GAUSS:1777", "BAYES:1702", "NEYMAN:1894", "LAPLACE:1749")
```

1.  Review: Split the names from the birth years using `str_split`, discarding the colon.  Assign the output to the variable `split_pioneers`.
```{r}
# Your code
```

2.  What class is the output object?
```{r}
# Your code
```

3.  Using `lapply`, convert the strings of `split_pioneers` to first-letter uppercase only.  Hint: `str_to_upper` is actually part of a family of functions.  Getting the documentation on it will reveal other functions like it.  Assign the output to the variable `statisticians`
```{r}
# Your code
```

4.  Try running `rbind` on `statisticians`.  What happens (it shouldn't be good)?
```{r}
# Your code
```
It creates a bizarre 1x4 matrix with vector elements.

5.  Using `do.call`, construct an `rbind` call that will turn the list into a matrix, `pioneers_df`
```{r}
# Your code
```

6.  Review: Convert the matrix into a data frame and name the columns `name` and `byear`.  Hint: `as...`
```{r}
# Your code
```

7.  Review: Convert the column types of `name` and `byear` to character and integer, respectively.  Hint: Subset and `as...` 
```{r}
# Your code
```

## Narcotics

The `sleep` dataset (built into R---no need to load anything) shows the effect of soporific (increases in sleep) drugs as compared to a control group.  There are two drugs represented in the data, as denoted by the group number.  The ID number is unique to each patient.

1.  Use `tapply` to find the average change in sleep hours for each of the drugs.
```{r}
# Your code
```

2.  Use `lapply` and an anonymous function to split the data frame into 10 2x3 sub-data frames grouped by ID.  Assign the output to `split_sleep`.  This is just practice thinking about `lapply` since you could actually just use this call in practice:
```{r, eval = F}
split(sleep, sleep$ID)
```

Hint: The first argument of `lapply` can be `1:10`.
```{r}
# Your code
```

3.  Use `vapply` on `split_sleep` to calculate the difference in extra sleep (group 2 - group 1) for each patient.  Hint: You can be a little fancy here and use `with` as your `FUN`.  Remember to give `vapply` a suitable template vector.
```{r}
# Your code
```

## Look Me in the Iris

Iris is a rather famous data set introduced by Ronald Fisher in his paper _The use of multiple measurements in taxonomic problems_ as an example of classification by dividing the data using (hyper)-planes.  It is already pre-loaded into `R` under `iris` (no package needed) and is often used as an example dataset in documentation.  I suggest inspecting the data set before continuing.

Use the data wrangling verbs of `dplyr` in conjunction with the piping operator `%>%` to do the following tasks.  Your output should be data frames assigned to the names in bold.

1.  Get only the sepal and petal length columns of the data frame.  __iris_len__
```{r}
# Your code
```
2.  Get every column of the data frame besides Petal length and width.  Do not call on the other three columns by their names.  __no_petal__
```{r}
# Your code
```
3.  Get only the species column, renaming it `species_name`.  __species__
```{r}
# Your code
```
4.  Create a column called `length_sum` that contains the sum of the petal and sepal lengths.  __add_length__
```{r}
# Your code
```
5.  Get only the rows of the data frame where the species is `setosa`. __only_setosa__
```{r}
# Your code
```
6.  Get only the rows of the data frame where the species name contains exactly 4 vowels and the sum of the sepal length and width is less than 8.  __nonsense_filter__
```{r}
# Your code
```
7.  Get the average sepal width `avg_sw` and average sepal length `avg_sl` for each of the 3 species.  __sepal_avg_by_species__
```{r}
# Your code
```
8.  Overwrite the `Sepal.Width` column so that it reads `"wide"` if the width is greater than 3 and `"narrow"` otherwise.  Do the same for `Sepal.Length` except the threshold is 5.5 and the replacement is `"long"` and `"short"`.  The `ifelse` command might be useful.  __factorized__
```{r}
# Your code
```
9.  Count how many of each wide-long, wide-short, narrow-long, and narrow-short pairs there are.  Hint: `group_by` followed by `tally` or `summarise`.  __sepal_counts__
```{r}
# Your code
```

