---
title: "lab25"
date: "July 26, 2016"
output: html_document
---

Write a function called `moveCars()` that moves the system one full step; that is, the function should move the red cars east once and the blue cars north once, with red cars going first.

Your function should take the following inputs:
`m`: a matrix

Your function should return a list containing two elements:
`m`: an updated matrix
`not_gridlocked`: a logical value that should be TRUE if the system changed and FALSE if not

```{r}

```


Check:
```{r}
m <- matrix(c(1,0,0,0,2,1,2,0,0,2,0,1,1,0,1,2,0,1,2,1,0,0,1,1,2), nrow = 5, ncol = 5, byrow = T)
identical(moveCars(m), list(matrix(c(0,1,0,0,2,1,2,0,0,2,1,1,0,1,0,2,0,1,2,1,0,0,1,1,2), nrow = 5, ncol = 5, byrow = T), TRUE))

m <-  matrix(c(0,1,0,2,1), nrow = 1, ncol = 5, byrow = T)
identical(moveCars(m), list(matrix(c(1,0,1,2,0), nrow = 1, ncol = 5, byrow = T), TRUE))

m <- matrix(c(2,0,1,2,0), nrow = 5, ncol = 1, byrow = T)
identical(moveCars(m), list(matrix(c(0,0,1,2,2), nrow = 5, ncol = 1, byrow = T), TRUE))
```

Now write a function called `initGrid()` that randomizes the initial system.

Your function should take the following inputs:
`r`: number of rows in the matrix
`c`: number of columns in the matrix
`d`: density of the grid 

Your function should return the following:
`m`: an rxc matrix with entries 0 (no cars), 1 (red cars), or 2 (blue cars)

```{r}

```






